### 2021-07-20 paper3/paper3.tex
Finish some stuff for paper3

We aim to convert our unlabelled data into a collection of \emph{perfect bags} \citep{kleinberg2016inherent,chouldechova2017fair}, \ie\ sample sets in which the class label $y$ and subgroup label $s$ are independent (\ie\ $y\perp s$).
Making use of the terminology from \emph{multiple-instance learning}, our batches comprise a certain number of bags which are a collection of samples.
%
We will then use these perfect bags as the inductive bias for learning the disentangled representations. 
%
How can we construct these perfect bags in the absence of any labelled data?
%
We assume that the number of subgroups is known \emph{a priori}\footnote{Relaxing this assumption represents a clear avenue for future work. We elaborate this in the limitation and intended use sec.~\ref{sec:limitations}.}. 
%We elaborate this further in our experiments and discussions sections.}.
%corresponding to the diverse demographic groups in the real-world population in which our machine learning system will be deployed. 
%
We then apply unsupervised k-means clustering, or a \emph{semi}-supervised clustering based on rank statistics; the latter allows incorporating annotations from the training data when forming the clusters.
%
Once the clusters have been found, we can sample from each cluster at an equal rate to form balanced (\ie\,\emph{perfect}) bags and use them as input for learning a disentangled representation. 
%
%See fig. \ref{fig:teaser} for an overview of our learning with invisible demographic framework.
We cast the problem of disentangling as one of distribution matching and propose an adversarial learning approach.
%a parametric approach based on adversarial learning. 
In the standard GAN setting, the discriminator assigns a score to each sample corresponding to the perceived probability that it was drawn from the true data distribution, and not the generator's.
In contrast, we train a discriminator to assign scores at the level of bags of samples, with a bag being deemed authentic if it was drawn from the originally unbiased distribution and not the de-biased one (with the de-biaser playing the role of the generator). To do so, we take inspiration from set-classification and multiple-instance learning and equip the discriminator with a learnable attention mechanism to model interdependencies between samples in a bag.

@@

@@

% \textbf{Disentangled representations via adversarial autoencoders}
%We want to learn a disentanglement using a perfect dataset as the supervision signal.
%
%To achieve this, our semi-supervised disentanglement learning framework (fig. \ref{fig:architecture} ) maps data points into independent factors of variations such that our training dataset $D_{\text{tr}}$ and our deployment set $D_c$ are indistinguishable; i.e., we would like transform $D_{\text{tr}}$ such that the independence condition $y\perp s$ holds.
%
%However, the deployment set itself $D_c$ might not amount to a suitable target distribution in which the independence condition holds.
%
%The idea is to make the deployment set $D_c$ approximately perfect by balancing it according to the cluster IDs, forming the set $D_p$.
%
% We create disentangled representations by training four separate neural networks, which we denote $f$, $g$, $h$ and $l$.
% %
% We use mean squared error for the reconstruction loss, and the binary cross entropy error for the discriminator loss.
% %
% For a dataset with mixed-typed attributes, we use a combination of mean squared error and (binary) cross entropy error for the reconstruction loss.
% %
% We optimise the proposed model using a scheduled update scheme where we freeze the weights of the autoencoder and predictor modules when we update the weights of the discriminator, and vice versa.
%
\section{Conclusion}
We have highlighted the problem that systematic bias can result in one or more subgroups having zero labelled data, and by doing so hope to have stimulated serious consideration for it (even if to be dismissed) when planning, building, evaluating and regulating machine learning systems.
We propose a two-step approach for addressing the resulting zero-label stratification problem.
First, we construct perfect bags from an unlabelled deployment set via  clustering. 
Second, we learn a disentangled representations using the perfect bags for adversarial distribution-matching. 
We empirically validate our framework on the Coloured MNIST, CelebA and Adult Income datasets, and find evidence that it is possible to maintain high performance on the subgroups with zero training data. 
We analyse our approach using the disentanglement calculus of \citet{ShuCheKumErmetal20} that relies on the notions of restrictiveness and consistency, and show that we can derive some guarantees.
Future work includes further restricting the amount of information that can be encoded in the bias factor, and the limitations that we mentioned in sec.~\ref{sec:limitations}.
Furthermore, the literature on clustering can potentially provide further insights on theoretical guarantees for clustering. This was not a focus in this paper as the disentangling aspect is more crucial to the approach.

>>>>

We aim to convert our unlabelled data into a collection of \emph{perfect bags} \citep{kleinberg2016inherent,chouldechova2017fair}, \ie\ sample sets in which the class label $y$ and subgroup label $s$ are independent (\ie\ $y\perp s$).
Making use of the terminology from \emph{multiple-instance learning}, our batches comprise a certain number of bags which are a collection of samples.
%
We will then use these perfect bags as the inductive bias for learning the disentangled representations. 
The disentangling procedure is thus in a sense \emph{supervised}.
%
How can we construct these perfect bags in the absence of any labelled data?
%
We assume that the number of subgroups is known \emph{a priori}\footnote{Relaxing this assumption represents a clear avenue for future work. We elaborate this in the limitation and intended use sec.~\ref{sec:limitations}.}. 
%We elaborate this further in our experiments and discussions sections.}.
%corresponding to the diverse demographic groups in the real-world population in which our machine learning system will be deployed. 
%
We then apply unsupervised k-means clustering, or a \emph{semi}-supervised clustering based on rank statistics; the latter allows incorporating annotations from the training data when forming the clusters.
%
Once the clusters have been found, we can sample from each cluster at an equal rate to form balanced (\ie\,\emph{perfect}) bags and use them as input for learning a disentangled representation. 
%
%See fig. \ref{fig:teaser} for an overview of our learning with invisible demographic framework.
We cast the problem of disentangling as one of distribution matching and propose an adversarial learning approach.
%a parametric approach based on adversarial learning. 
In the standard GAN setting, the discriminator assigns a score to each sample corresponding to the perceived probability that it was drawn from the true data distribution, and not the generator's.
In contrast, we train a discriminator to assign scores at the level of bags of samples, with a bag being deemed authentic if it was drawn from the originally unbiased distribution and not the de-biased one (with the de-biaser playing the role of the generator). To do so, we take inspiration from set-classification and multiple-instance learning and equip the discriminator with a learnable attention mechanism to model interdependencies between samples in a bag.

@@

Note that as long as the model sees the different \(s\)-\(y\)-combinations at the right proportions,
disentangling can (to an extent) also be achieved without the bag-wise loss
(\ie, when the bag size is set to 1).
The model then has to learn an implicit prior for which \(s\)-\(y\)-combinations to expect
in the two different sets -- training set and deployment set.
However, in our preliminary experiments we found this sample-wise approach to work much poorer
than using a bag-wise loss.
For those experiments, the batch was balanced in the same way as described above for the bags,
but the batch was not subdivided into bags
and there was no attention or aggregation method applied to the batch:
the loss was just computed per sample.
From the poor performance,
we inferred that the sample-wise loss lacks the necessary context
to easily differentiate between the two datasets.

Our contribution to the disentanglement problem is thus two-fold:
i) the use of the difference between two distinct sets as the supervision signal,
and ii) the use of a bag-wise loss
(where the bags are sampled to represent their corresponding data distribution in an idealised way)
which allows the model to contrast the two sets more easily.
Together, they form the idea of disentangling with ``perfect bags''.

@@

% \textbf{Disentangled representations via adversarial autoencoders}
%We want to learn a disentanglement using a perfect dataset as the supervision signal.
%
%To achieve this, our semi-supervised disentanglement learning framework (fig. \ref{fig:architecture} ) maps data points into independent factors of variations such that our training dataset $D_{\text{tr}}$ and our deployment set $D_c$ are indistinguishable; i.e., we would like transform $D_{\text{tr}}$ such that the independence condition $y\perp s$ holds.
%
%However, the deployment set itself $D_c$ might not amount to a suitable target distribution in which the independence condition holds.
%
%The idea is to make the deployment set $D_c$ approximately perfect by balancing it according to the cluster IDs, forming the set $D_p$.
%
% We create disentangled representations by training four separate neural networks, which we denote $f$, $g$, $h$ and $l$.
% %
% We use mean squared error for the reconstruction loss, and the binary cross entropy error for the discriminator loss.
% %
% For a dataset with mixed-typed attributes, we use a combination of mean squared error and (binary) cross entropy error for the reconstruction loss.
% %
% We optimise the proposed model using a scheduled update scheme where we freeze the weights of the autoencoder and predictor modules when we update the weights of the discriminator, and vice versa.
%
\section{Conclusion}
We have highlighted the problem that systematic bias can result in one or more subgroups having zero labelled data, and by doing so hope to have stimulated serious consideration for it (even if to be dismissed) when planning, building, evaluating and regulating machine learning systems.
We propose a two-step approach for addressing the resulting zero-label stratification problem.
First, we construct perfect bags from an unlabelled deployment set via  clustering. 
Second, we learn a disentangled representations using the perfect bags for adversarial distribution-matching. 
We empirically validate our framework on the Coloured MNIST, CelebA and Adult Income datasets, and find evidence that it is possible to maintain high performance on the subgroups with zero training data. 
We analyse our approach using the disentanglement calculus of \citet{ShuCheKumErmetal20} that relies on the notions of restrictiveness and consistency, and show that we can derive some guarantees.
We presented our approach in the context of biased data,
but it could be used as a general-purpose method for learning disentanglements
as long as the factors can be expressed by contrasting two data subsets;
that is, the data will be disentangled into two factors,
the first of which corresponding to the aspect that is common in the two sets,
and the second corresponding to the one that differs.
The method thus does not compete directly with \emph{unsupervised} disentanglement,
but is potentially more widely applicable than to only fairness problems.

Future work includes further restricting the amount of information that can be encoded in the bias factor, and the limitations that we mentioned in sec.~\ref{sec:limitations}.
Furthermore, the literature on clustering can potentially provide further insights on theoretical guarantees for clustering. This was not a focus in this paper as the disentangling aspect is more crucial to the approach.
